Shifting paradigms in Ocean Color: Bayesian Inference for Uncertainty-Aware Chlorophyll Estimation 

### 2025-05-18 manuscript/manuscript.qmd
updated methods section  intro

---
title: "Shifting paradigms in Ocean Color: Bayesian Inference for Uncertainty-Aware Chlorophyll Estimation "
format:
    agu-pdf:
        keep-tex: true
    agu-html: default
author:  
  - name: Erdem M. Karaköylü
    orcid: 0000-0002-6156-1720
    corresponding: true
    email: erdemk@protonmail.com
    roles:
      - Investigator
      - Software
      - Visualization
      - Project administration
    affiliations:
      - Independant Researcher
  - name: Susanne E. Craig
    orcid: 0000-0002-8963-0951
    corresponding: false
    email: susanne.e.craig@nasa.gov
    roles:
      - Co-Investigator
    affiliations:
      - NASA/UMBC
abstract: Placeholder
plain-language-summary: Placeholder
keywords: ['Bayesian modeling', 'Ocean Color', ]
key-points:
  - Bayesian models have not been fully adopted in Ocean Color remote sensing yet.
  - A Bayesian workflow is a principled guide to a fully Bayesian study
  - 
bibliography: references.bib
citation:
  container-title: Geophysical Research Letters
keep-tex: true
number-sections: false
date: last-modified
---

@@

I used the NOMAD v2 (2008) dataset, which includes quality-controlled in situ measurements of chlorophyll-a matched with coincident satellite Rrs data. These measurements span a range of oceanographic conditions, enabling the development of models that generalize across water types.

@@

This section presents convergence diagnostics, posterior inference, and model comparison using fully Bayesian metrics. No frequentist evaluation criteria were used. Instead, we rely on information-theoretic model comparison (LOO-CV), posterior predictive checks, and visualization of uncertainty via highest density intervals (HDIs).

>>>>

---
title: "Shifting paradigms in Ocean Color: Bayesian Inference for Uncertainty-Aware Chlorophyll Estimation "
format:
    agu-pdf:
        keep-tex: true
    agu-html: default
author:  
  - name: Erdem M. Karaköylü
    orcid: 0000-0002-6156-1720
    corresponding: true
    email: erdemk@protonmail.com
    roles:
      - Investigator
      - Software
      - Visualization
      - Project administration
    affiliations:
      - Independant Researcher
  - name: Susanne E. Craig
    orcid: 0000-0002-8963-0951
    corresponding: false
    email: susanne.e.craig@nasa.gov
    roles:
      - Co-Investigator
    affiliations:
      - NASA/UMBC
abstract: Placeholder
plain-language-summary: Placeholder
keywords: ['Bayesian modeling', 'Ocean Color', ]
key-points:
  - Bayesian models have not been fully adopted in Ocean Color remote sensing yet.
  - A Bayesian workflow is a principled guide to a fully Bayesian study
  - 
bibliography: references.bib
citation:
  container-title: Geophysical Research Letters
keep-tex: true
number-sections: true
date: last-modified
---

@@

For pedagogical reasons, I used the well known NOMAD [@werdell2005improved] dataset, which includes quality-controlled in situ measurements of chlorophyll-a matched with coincident satellite Rrs data. These measurements span a range of oceanographic conditions, enabling the development of models that generalize across water types. Sea-surface reflectances (Rrs) in the 6 visible SeaWiFS bands (411, 443, 489, 510, 550, 670nm) constitue the bulk of the data. As prediction targets corresponding chlorophyll concentrations using both fluorescence and HPLC were retained.

@@

This section presents the outputs of posterior inference, model comparison using Bayesian criteria, and posterior predictive diagnostics for the four primary models. Results from the three supplementary models are provided separately in the Supplement.
